Algebraic File Synchronization: Adequacy and Completeness

### 2017-11-23 p2/axioms.tex
Updated script

It is easy to see that this equivalence holds on the level of filesystems:

\begin{mylem}\mlabel{equiv_on_fs}
$A\equiv B$
iff $A$ and $B$ behave in the same way on
all filesystems, that is, $\forall \FS: A\aFS=B\aFS$.
\end{mylem}
\begin{proof}
% TODO APPENDIX
If $\forall \FS: A\aFS=B\aFS$, then $A\eqext B$ and $B\eqext A$ and so by definition $A\equiv B$.
For the reverse statement we know $A\eqext B$ and $B\eqext A$.
$B\eqext A$ implies that it is not possible for both $B\aFS\neq\fsbroken$ and $A\aFS=\fsbroken$ to be true.
Therefore, for every filesystem $\FS$, either $A\aFS\neq\fsbroken$, in which case $A\aFS=B\aFS$ as $A\eqext B$, or $A\aFS=\fsbroken$,
in which case $B\aFS=\fsbroken$ must be true, and therefore $A\aFS=B\aFS$.
\end{proof}

@@

Our aim is to derive information about the effects of sequences
of commands independently of the actual filesystems.
In order to make this possible, we investigate the smallest building
blocks of sequences: pairs of commands that act on a filesystem directly one after the other.
This approach is useful as there are a limited number of command pairs,
because, as we argued above, we can disregard the exact output values of commands apart from their type,
and we can also abstract the relationship between the nodes in the two commands
to a finite number of cases.

We methodically investigate all possibilities using a computer 
program\footnote{The program is accessible on-line at \\
https://github.com/csirmaz/AlgebraicSyncPaper/blob/master/p2/prove.py.}
to determine
which pairs of commands cause errors all the time,
which can be simplified to one command or $\emptyseq$,
and which commute and can be reversed without any change in their overall effect.
Our basis for this investigation was the model of filesystems introduced in this paper.
These properties of command pairs are crucial as they determine
how a set of commands can be re-ordered to be applied to a filesystem
during synchronization, and what commands will never be compatible.
Below we list a number of statements derived using this method,
which are in fact lemmas that can be derived from our definitions of filesystems and of commands.

@@

\bigskip

\noindent
Pairs of commands in general have the form
\[ \cxynv\cc  \czwmv \]
where $x,y,z,w\in\setv$ and $n,m\in\setn$.

@@

\bigskip

@@

\noindent
We also define the concept of two commands being independent:

>>>>

It is easy to see that the equivalence $\equiv$ holds on the level of filesystems:
\begin{mylem}\mlabel{equiv_on_fs}
$A\equiv B$
iff $A$ and $B$ behave in the same way on
all filesystems, that is, $\forall \FS: A\aFS=B\aFS$.
\end{mylem}
\begin{proof}
As $A\equiv B$ means that both $A$ extends $B$ and $B$ extends $A$, it necessarily
follows that $A$ and $B$, as partial functions, are identical.
\end{proof}

@@

\myskip
Our aim is to derive information about the effects of sequences
of commands independently of the actual filesystems.
In order to make this possible, we investigate the smallest building
blocks of sequences: pairs of commands that act on a filesystem directly one after the other.
This approach is useful as there are a limited number of command pairs,
because, as we argued above, we can disregard the exact output values of commands apart from their type,
and we can also abstract the relationship between the nodes in the two commands
to a finite number of cases.
These properties of command pairs are crucial as they determine
how a set of commands can be re-ordered to be applied to a filesystem
during synchronization, and what command pairs will never be compatible.

% TODO \footnote{The program is accessible on-line at \\
% https://github.com/csirmaz/AlgebraicSyncPaper/blob/master/p2/prove.py.}

@@

\myskip
Command pairs in genral have the form
\[ \cxynv\cc  \czwmv \]
where $x,y,z,w\in\setv$ and $n,m\in\setn$.

@@

@@

\myskip
We also define the concept of two commands being independent:

### 2017-11-23 p2/def-commands.tex
Updated script

\bigskip
\noindent
Depending on their input and output types, there are nine types of commands,
which we separate into four categories depending on the overall effect of the command.
Constructions commands extend the filesystem,
while destruction commands shrink it;
the replacement command replaces a file value with a (presumably different) file value,
and assertion commands either break the filesystem, or assert the type of the value at a node.
\begin{mydef}[Command categories]
Depending on their input and output types, filesystem commands
belong to exactly one of the following four categories:
\begin{itemize}
\item[]\emph{Construction commands:}
    $\cbfa{n}$, $\cbda{n}$ and $\cfda{n}$;
\item[]\emph{Destruction commands:}
    $\cdfa{n}$, $\cdba{n}$ and $\cfba{n}$;
\item[]\emph{Assertion commands:}
    $\cbba{n}$ and $\cdda{n}$;
\item[]\emph{Replacement command:}
    $\caaa{\valfx}{\valff}{n}$.
\qedhere
\end{itemize}
\end{mydef}

@@

\bigskip

\noindent
For reasons also listed in \cite{NREC}, in this paper we will not consider
a \textit{move} or \textit{rename} command.
This turns out to be useful because this would be the only command that affects
filesystems at two nodes at once, therefore describing 
the dependencies for \textit{move} would call for a more complicated model.
From a pragmatic perspective, this restriction does not mean that in an application
implementing conflict resolution using the algorithm described here would not be
able to handle renames by pre- and post-processing changes in the filesystem to
discover them, which can enhance usability, but which
(especially when a rename is combined with changes to the content)
is a non-trivial problem in itself.

>>>>

\myskip
Depending on their input and output types, there are nine types of commands,
which we separate into four categories depending on the overall effect of the command.
Constructions commands extend the filesystem,
while destruction commands shrink it;
the replacement command replaces a file value with a (presumably different) file value,
and assertion commands either break the filesystem, or assert the type of the value at a node.
\begin{mydef}[Command categories]
Depending on their input and output types, filesystem commands
belong to exactly one of the following four categories:
\begin{itemize}
\item[]\emph{Construction commands:}
    $\cbfa{n}$, $\cbda{n}$ and $\cfda{n}$;
\item[]\emph{Destruction commands:}
    $\cdfa{n}$, $\cdba{n}$ and $\cfba{n}$;
\item[]\emph{Assertion commands:}
    $\cbba{n}$ and $\cdda{n}$;
\item[]\emph{Replacement command:}
    $\caaa{\valfx}{\valff}{n}$.
\qedhere
\end{itemize}
\end{mydef}

@@

\myskip
For reasons also listed in \cite{NREC}, in this paper we will not consider
a \textit{move} or \textit{rename} command.
This turns out to be useful because this would be the only command that affects
filesystems at two nodes at once, therefore describing 
the dependencies for \textit{move} would call for a more complicated model.
From a pragmatic perspective, this restriction does not mean that in an application
implementing conflict resolution using the algorithm described here would not be
able to handle renames by pre- and post-processing changes in the filesystem to
discover them, which can enhance usability, but which
(especially when a rename is combined with changes to the content)
is a non-trivial problem in itself.

### 2017-11-23 p2/paper.tex
Updated script

\bigskip\noindent
{\bf Keywords:} % (4-6)
file synchronization,
algebraic approach,
distributed systems,
reconciliation,
data replication
% conflict detection,
% conflict resolution
% computer algebra

\bigskip\noindent
{\bf MSC classes:} % http://www.ams.org/mathscinet/msc/msc2010.html
08A70, % Applications of universal algebra in computer science
68M07, % Mathematical problems of computer architecture
68M14, % Distributed systems
68P05, % Data structures
68P20 % Information storage and retrieval
% 68N25         Operating systems

\bigskip\noindent
{\bf ACM classes:} % http://www.acm.org/about/class/ccs98-html
% (D) Software (D.4) Operating systems
D.4.3, % File systems management
% (E) Data
E.5, % Files
% (F) Theory of computation (F.2) ANALYSIS OF ALGORITHMS AND PROBLEM COMPLEXITY
F.2.2, % Nonnumerical Algorithms and Problems
% (G) Mathematics of Computing
G.2 % DISCRETE MATHEMATICS
% (H) Information Systems (H.3) INFORMATION STORAGE AND RETRIEVAL
% H.3.4 % Systems and Software

>>>>

\myskip
{\bf Keywords:} % (4-6)
file synchronization,
algebraic approach,
distributed systems,
reconciliation,
data replication
% conflict detection,
% conflict resolution
% computer algebra

\myskip
{\bf MSC classes:} % http://www.ams.org/mathscinet/msc/msc2010.html
08A70, % Applications of universal algebra in computer science
68M07, % Mathematical problems of computer architecture
68M14, % Distributed systems
68P05, % Data structures
68P20 % Information storage and retrieval
% 68N25         Operating systems

\myskip
{\bf ACM classes:} % http://www.acm.org/about/class/ccs98-html
% (D) Software (D.4) Operating systems
D.4.3, % File systems management
% (E) Data
E.5, % Files
% (F) Theory of computation (F.2) ANALYSIS OF ALGORITHMS AND PROBLEM COMPLEXITY
F.2.2, % Nonnumerical Algorithms and Problems
% (G) Mathematics of Computing
G.2 % DISCRETE MATHEMATICS
% (H) Information Systems (H.3) INFORMATION STORAGE AND RETRIEVAL
% H.3.4 % Systems and Software
